---
title: "PA1_template.Rmd"
author: "Mohau Mateyisi"
date: "23 January 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

```{r}
library(data.table);
library(lubridate);
library(plyr)

Adata <- read.csv('file:///D:/Cousera/reproducible-research/activity.csv')
Adata$date<- dmy(Adata$date)

```


## What is mean total number of steps taken per day?



```{r}
dt2<-aggregate(steps ~ date , data = na.exclude(Adata), sum) 

hist(dt2$steps,
     main="Histogram of Total number of steps per day", 
     xlab="steps", 
     border="blue", 
     col="green")




names(dt2)<-c("date","TtStepsPerDay")
ss<-summary(dt2)

minim  <-unlist(strsplit(ss[,"TtStepsPerDay"][1],":"))[2]
mediann <-unlist(strsplit(ss[,"TtStepsPerDay"][3],":"))[2]
meanv <-unlist(strsplit(ss[,"TtStepsPerDay"][4],":"))[2]

```

The mean and median number of steps are `r meanv` and `r mediann` respectively. 

## What is the average daily activity pattern?

Below we plot the average daily activity pattern. 

```{r,echo=TRUE}
library(ggplot2)
dtt2<-aggregate(steps ~ interval , data = na.exclude(Adata), mean) 

with(dtt2, {  
  plot(interval, steps, type="l", xlab="inerval", ylab="Av_steps", main="Average daily activity pattern", col="red")   
})


maximStepsInterval <-max(dtt2$interval)

# mean and median of the total number of steps per day.

meann1<-mean(dtt2$steps)
median1<- median(dtt2$steps)


```

On average the maximaum happen on the `r maximStepsInterval`th 5 minute interval.

The mean and median of the total number of steps per day are `r meann1 ` and `r median1` respectively


## Imputing missing values

The imputation method aggregate the number of steps for each day and fills the missing value
for each day with the mean. In the case where all the values per day are missing the imputation leaves value for the day empty.

```{r}


library("imputeMissings")
Adata1<-Adata[,c("date","steps","interval")]

#impute with the median
r<-impute(Adata1, object = NULL, method = "median/mode", flag = FALSE)
zz<-summary(Adata)


dtt22<-aggregate(steps ~ date , data = na.exclude(r), sum) 

hist(dtt22$steps, 
     main="Histogram of Total number of steps per day after imputation", 
     xlab="steps", 
     border="blue", 
     col="green")


ismissing <-unlist(strsplit(zz[[7]],":"))[2]


dt22<-aggregate(steps ~ date , data = r, sum) 

zz2 <- summary(dt22)

mmean<- unlist(strsplit(zz2[4,][2],":"))[2]
mmedian <-unlist(strsplit(zz2[3,][2],":"))[2]

```

There are about `r ismissing` missing values.

The mean and median after data amputation are `r mmean`  and `r mmedian`.

The imputation process leads to a lower mean and median compared that obtained before replacing NA's.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
library("ggplot2")
library("prettyR")



dt3<-r
dt3$weekday<-weekdays(r$date)

df1 = subset(dt3,dt3$weekday == "Saturday" | dt3$weekday == "Sunday")
df1<-aggregate(steps ~ interval , data = na.exclude(df1), mean)
df1$day<- "weekend"


df2 = subset(dt3,dt3$weekday != "Saturday" & dt3$weekday != "Sunday")
df2<-aggregate(steps ~ interval , data = na.exclude(df2), mean)
df2$day<-"weekday"


df3<-rbind(df1,df2)




p1 <-ggplot(df3, aes(interval,steps ))
p1 + geom_line(stat = "identity")+
   facet_wrap( ~ day, nrow = 2) + theme(legend.position = "none") +
   labs(title = "Comparing weekend and week day activities")+
   theme(text = element_text(size=20),
        axis.text.x = element_text(angle=90, hjust=1))



```

Between the 500 and 1000 time interval week day steps are much higher then the weekend steps. However, after the 1000 interval 5 min the average steps during the week are relatively lower than the average steps during the weekend.


